# Reproducible Research: Peer Assessment 1
##########################################

## Loading and preprocessing the data 
#######################################
*1. Load the data*
```{r}
acti_data <- read.csv("activity.csv")
summary(acti_data)
head(acti_data)
```
#######################################################################################

## What is mean total number of steps taken per day?

*1.Calculate the total number of steps taken per day*
```{r}
steps <- aggregate(acti_data$steps, by = list(Date = acti_data$date), FUN = sum)
library(ggplot2)
names(steps)[names(steps) == "x"] <- "Total"
temp <- as.Date(steps$Date, "%Y-%m-%d")
steps$Date <- format(temp, format = "%m-%d")
head(steps)
```

*2. Make a histogram of the total number of steps taken each day*
```{r}
hist1 <- ggplot(data = na.omit(steps), aes(Total)) + 
    geom_histogram(binwidth = 1500, colour = "white") +
    xlab("Total Number of Steps Taken Each Day") +
    ylab("Count") +
    ggtitle("Histogram of the Total Number of Steps Taken Each Day")
print(hist1)
```

*3. Calculate and report the mean and median total number of steps taken per day*
```{r}
mean(na.omit(steps$Total))
median(na.omit(steps$Total))
```
########################################################################################

## What is the average daily activity pattern?

*1. Make a time series plot*
```{r}
five_min_steps <- aggregate(steps ~ interval, data = acti_data, FUN =mean)
TimeSeries1 <- ggplot(data = five_min_steps, aes(x = interval, y = steps)) + 
    geom_line() +
    xlab("Time Intervals (5 Minutes is an unit)") + 
    ylab("Total Number of Steps") +
    ggtitle("Average Number of Steps Taken of the 5-Minute Interval")
print(TimeSeries1) 
```

*2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?*
```{r}
head(five_min_steps)
five_min_steps[which(five_min_steps$steps == max(five_min_steps$steps)),]
```
###########################################################################################

## Imputing missing values

*1. Calculate and report the total number of missing values in the dataset*
```{r}
sapply(X = acti_data, FUN = function(x) sum(is.na(x)))
```


*2. Devise a strategy for filling in all of the missing values in the dataset.*
```{r}
library(dplyr)
replace_with_mean <- function(num) replace(num, is.na(num), mean(num, na.rm = TRUE))
meanday <- (acti_data %>% group_by(interval) %>% mutate(steps = replace_with_mean(steps)))
head(meanday)
sum(is.na(meanday))
```

*3. Create a new dataset that is equal to the original dataset but with the missing data filled in*
```{r}
new_dataset <- as.data.frame(meanday)
head(new_dataset)
summary(new_dataset)
```

*4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.*
```{r}
new_steps <- aggregate(new_dataset$steps, by = list(new_dataset$date), FUN = sum)
names(new_steps)[names(new_steps) == "x"] <- "Total"
names(new_steps)[names(new_steps) == "Group.1"] <- "Date"
hist2 <- ggplot(data = new_steps, aes(Total)) + 
    geom_histogram(binwidth = 1500, colour = "white") +
    xlab("Total Number of Steps Taken Each Day") +
    ylab("Count") +
    ggtitle("Histogram of the Total Number of Steps Taken Each Day with New Version Dataset")
print(hist2)
```

Now, we compare two plots.
```{r}
library(grid)
library(gridExtra)
grid.arrange(hist1, hist2, ncol = 2)
```

Compare the mean and median.
```{r}
mean(na.omit(steps$Total))
median(na.omit(steps$Total))
mean(new_steps$Total)
median(new_steps$Total)
```
###########################################################################################

## Are there differences in activity patterns between weekdays and weekends?

*1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.*
```{r}
new_dataset$WeekendOrWeekday <- ifelse(weekdays(as.Date(new_dataset$date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "Weekday", "Weekend")
head(new_dataset)
```

*2. Make a panel plot containing a time series plot*
```{r}
new_dataset <- (new_dataset %>% group_by(interval, WeekendOrWeekday) %>% summarise(Mean = mean(steps)))
ggplot(new_dataset, mapping = aes(x = interval, y = Mean)) + geom_line() +
    facet_grid(WeekendOrWeekday ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval")
```




























